\pkg{Rgbp}: An R Package for Gaussian, Poisson, and Binomial Hierarchical Modeling

### 2014-04-21 jss/article.tex
paper and graphs

\begin{CodeChunk}
\begin{CodeInput}
R> z <- c( 3,   2,   5,  11,   9,  12,  12,   4,  10,  13,  14,   7,  12,
          11,  13,  22,  15,  11,  14,  11,  16,  14,   9,  15,  13,  35,
          26,  25,  20,  35,  27)
R> n <- c(67,  68, 210, 256, 269, 274, 278, 295, 347, 349, 358, 396, 431,
         441, 477, 484, 494, 501, 505, 540, 563, 593, 602, 629, 636, 729,
         849, 914, 940, 1193, 1340)
\end{CodeInput}
\end{CodeChunk}
or
\begin{CodeChunk}
\begin{CodeInput}
R> data("hospital")
R> z <- hospital$d # variable name for the number of deaths is d in dataset
R> n <- hospital$n
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> plot(p)
\end{CodeInput}
\end{CodeChunk}
\begin{figure}[h]
\begin{center}
\includegraphics[width = 6in]{hospital1.png}
\caption{Shrinkage plot and 95\% interval plot for 31 hospitals}
\label{fig:hospshr}
\end{center}
\end{figure}

@@

In Figure \ref{fig:hospitalcoverage}, the black horizontal line at 0.95 represents the nominal confidence level and the smooth red curve connects Rao-Blackwellized (RB) unbiased coverage estimates for 31 hospitals. The overall RB unbiased coverage estimate across all the hospitals is 0.954. And none of RB unbiased coverage estimates for 31 hospitals are less than 0.95 regardless of their caseloads ($n_{j}$). This result shows that the interval estimates for this particular dataset accurately achieves 95\% confidence under repeated sampling. Note that these estimates depend on the given true value of $r$, the known prior mean, and the assumption that the model is true.

The following code provides 31 RB unbiased coverage estimates for each hospital.
\begin{CodeChunk}
\begin{CodeInput}
R> pcv$coverageRB
\end{CodeInput}
\begin{CodeOutput}
 [1] 0.956 0.956 0.956 0.955 0.955 0.955 0.955 0.955 0.955 0.955 0.955 0.954 
[13] 0.955 0.954 0.954 0.954 0.954 0.954 0.955 0.954 0.954 0.953 0.953 0.953
[25] 0.954 0.954 0.954 0.953 0.953 0.953 0.952
\end{CodeOutput}
\end{CodeChunk}

And the code below shows 31 simple unbiased coverage estimates for each hospital.
\begin{CodeChunk}
\begin{CodeInput}
R> pcv$coverageS
\end{CodeInput}
\begin{CodeOutput}
 [1] 0.947 0.952 0.963 0.961 0.957 0.957 0.959 0.952 0.957 0.957 0.941 0.958 
[13] 0.954 0.962 0.935 0.940 0.955 0.952 0.951 0.941 0.955 0.952 0.954 0.946 
[25] 0.942 0.952 0.958 0.945 0.938 0.958 0.948
\end{CodeOutput}
\end{CodeChunk}

The function \code{coverage} also calculates the standard errors for each hospital's RB unbiased coverage estimate defined in (\ref{RBvar}). The following code provides 31 standard errors for RB estimates.
\begin{CodeChunk}
\begin{CodeInput}
R> pcv$se.coverageRB
\end{CodeInput}
\begin{CodeOutput}
 [1] 0.0017 0.0017 0.0014 0.0014 0.0014 0.0014 0.0014 0.0014 0.0013 0.0013 0.0013 
[12] 0.0013 0.0012 0.0012 0.0012 0.0012 0.0011 0.0012 0.0011 0.0011 0.0011 0.0011
[23] 0.0010 0.0011 0.0010 0.0010 0.0009 0.0009 0.0009 0.0007 0.0007
\end{CodeOutput}
\end{CodeChunk}
Similarly, 31 standard errors for each simple unbiased coverage estimate defined in (\ref{svar}) are
\begin{CodeChunk}
\begin{CodeInput}
R> pcv$se.coverageS
\end{CodeInput}
\begin{CodeOutput}
 [1] 0.0071 0.0068 0.0060 0.0061 0.0064 0.0064 0.0063 0.0068 0.0064 0.0064 0.0075 
[12] 0.0063 0.0066 0.0060 0.0078 0.0075 0.0066 0.0068 0.0068 0.0075 0.0066 0.0068 
[23] 0.0066 0.0072 0.0074 0.0068 0.0063 0.0072 0.0076 0.0063 0.0070
\end{CodeOutput}
\end{CodeChunk}

Taking the first hospital as an example, the variance estimate of RB unbiased coverage estimate is about 17 times smaller than that of simple one. It means that 1,000 RB unbiased coverage estimates are as precise as 17,000 simple unbiased coverage estimates in terms of estimating true coverage probability for the first hospital, $p_{cov, 1}$.

@@

The Education Testing Service (ETS) conducted randomized experiments in eight separate schools (groups) to test whether students (units) SAT scores are effected by coaching. The dataset contains the estimated coaching effects on SAT scores ($y_{j}, j=1, \ldots, 8$) and standard errors ($se_{j}, j=1, \ldots, 8$) of the eight schools \citep{1981}.
\begin{CodeChunk}
\begin{CodeInput}
R> y  <- c(12, -3, 28,  7,  1,  8, 18, -1)
R> se <- c(18, 16, 15, 11, 11, 10, 10,  9)
\end{CodeInput}
\end{CodeChunk}
or
\begin{CodeChunk}
\begin{CodeInput}
R> data("schools")
R> y  <- schools$y
R> se <- schools$se
\end{CodeInput}
\end{CodeChunk}

@@

\begin{figure}[h] 
\begin{center}
\includegraphics[width = 6in]{school1.png}
\caption{Shrinkage plot and 95\% interval plot for 8 schools}
\label{fig:8schoolsplot}
\end{center}
\end{figure}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> z <- c(18, 17, 16, 15, 14, 14, 13, 12, 11, 11, 10, 10, 10, 10, 10,  9,  8,  7)
R> n <- c(45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45)
R> x <- c( 1,  1,  1,  1,  1,  0,  0,  0,  0,  1,  0,  0,  0,  1,  1,  0,  0,  0) 
\end{CodeInput}
\end{CodeChunk}
or
\begin{CodeChunk}
\begin{CodeInput}
R> data("baseball")
R> z <- baseball$Hits
R> n <- baseball$At.Bats
R> x <- ifelse(baseball$Position == "fielder", 1, 0)
\end{CodeInput}
\end{CodeChunk}
%$
Conditioning on the true batting average for each player we assume that the at-bats are independent and therefore, $z_{j}\vert p_{j}\stackrel{ind}{\sim} \textrm{Binomial}(45, p_{j}), ~j=1, \ldots, 18$. Our goal is to obtain point and interval estimates of the true batting average, $p_{j}$, for each player, whilst considering the additional information on whether the player is an outfielder or not. \code{gbp} provides a way to incorporate such covariate information seamlessly into the second-level hierarchy such that information is shared and regression towards the mean (RTTM) occurs within outfielders and non-outfielders separately. %For reference, we can also assume the Normal distribution using $y_{j}=z_{j} / n_{j}$ and $V_{j}=\bar{y}(1-\bar{y})/n$, where $\bar{y}=\sum_{j} z_{j} / \sum_{j} n_{j}$, $j=1,\ldots, 18$.

@@

\begin{CodeChunk}
\begin{CodeInput}
R> plot(b)
\end{CodeInput}
\end{CodeChunk}
\begin{figure}[h]
\begin{center}
\includegraphics[width = 6in]{baseball1.png}
\caption{Shrinkage plot and 95\% interval plot for 18 baseball players}
\label{fig:baseball}
\end{center}
\end{figure}

@@

The Rao-Blackwellized unbiased coverage estimates and their standard errors for each player follow.
\begin{CodeChunk}
\begin{CodeInput}
R> bcv$coverageRB
\end{CodeInput}
%$
\begin{CodeOutput}
 [1] 0.972 0.973 0.973 0.973 0.972 0.972 0.972 0.972 0.973 0.972 0.973 0.972 
[13] 0.973 0.973 0.971 0.973 0.971 0.973
\end{CodeOutput}
\end{CodeChunk}
\begin{CodeChunk}
\begin{CodeInput}
R> bcv$se.coverageRB
\end{CodeInput}
\begin{CodeOutput}
 [1] 0.0015 0.0011 0.0012 0.0011 0.0015 0.0012 0.0012 0.0010 0.0011 0.0012 
[11] 0.0010 0.0012 0.0010 0.0011 0.0013 0.0011 0.0013 0.0010
\end{CodeOutput}
\end{CodeChunk}

>>>>

\begin{CodeChunk}
\begin{CodeInput}
R> z <- c( 3,   2,   5,  11,   9,  12,  12,   4,  10,  13,  14,   7,  12,
          11,  13,  22,  15,  11,  14,  11,  16,  14,   9,  15,  13,  35,
          26,  25,  20,  35,  27)
R> n <- c(67,  68, 210, 256, 269, 274, 278, 295, 347, 349, 358, 396, 431,
         441, 477, 484, 494, 501, 505, 540, 563, 593, 602, 629, 636, 729,
         849, 914, 940, 1193, 1340)
\end{CodeInput}
\end{CodeChunk}
or
\begin{CodeChunk}
\begin{CodeInput}
R> data(hospital)
R> z <- hospital$d # variable name for the number of deaths is d in dataset
R> n <- hospital$n
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> plot(p)
\end{CodeInput}
\end{CodeChunk}
\begin{figure}[h]
\begin{center}
\includegraphics[width = 3.2in]{hospital1.png}
\caption{Shrinkage plot and 95\% interval plot for 31 hospitals}
\label{fig:hospshr}
\end{center}
\end{figure}

@@

In Figure \ref{fig:hospitalcoverage}, the black horizontal line at 0.95 represents the nominal confidence level and the smooth red curve connects Rao-Blackwellized (RB) unbiased coverage estimates for 31 hospitals. The overall RB unbiased coverage estimate across all the hospitals is 0.953. And none of RB unbiased coverage estimates for 31 hospitals are less than 0.95 regardless of their caseloads ($n_{j}$). This result shows that the interval estimates for this particular dataset accurately achieves 95\% confidence under repeated sampling. Note that these estimates depend on the given true value of $r$, the known prior mean, and the assumption that the model is true.

The following code provides 31 RB unbiased coverage estimates for each hospital.
\begin{CodeChunk}
\begin{CodeInput}
R> pcv$coverageRB
\end{CodeInput}
\begin{CodeOutput}
 [1] 0.955 0.954 0.955 0.954 0.954 0.954 0.954 0.954 0.954 0.954 0.954 0.954
[13] 0.954 0.953 0.953 0.954 0.953 0.954 0.954 0.953 0.953 0.953 0.954 0.953
[25] 0.953 0.953 0.952 0.953 0.952 0.952 0.952
\end{CodeOutput}
\end{CodeChunk}

And the code below shows 31 simple unbiased coverage estimates for each hospital.
\begin{CodeChunk}
\begin{CodeInput}
R> pcv$coverageS
\end{CodeInput}
\begin{CodeOutput}
 [1] 0.949 0.960 0.958 0.958 0.950 0.945 0.960 0.953 0.960 0.956 0.955 0.946 
[13] 0.955 0.954 0.960 0.965 0.955 0.952 0.960 0.956 0.959 0.955 0.964 0.960 
[25] 0.945 0.942 0.947 0.960 0.940 0.946 0.956
\end{CodeOutput}
\end{CodeChunk}

The function \code{coverage} also calculates the standard errors for each hospital's RB unbiased coverage estimate defined in (\ref{RBvar}). The following code provides 31 standard errors for RB estimates.
\begin{CodeChunk}
\begin{CodeInput}
R> pcv$se.coverageRB
\end{CodeInput}
\begin{CodeOutput}
 [1] 0.0016 0.0016 0.0014 0.0014 0.0013 0.0013 0.0013 0.0013 0.0012 0.0013 0.0012 
[12] 0.0012 0.0011 0.0011 0.00110.0011 0.0011 0.0011 0.0011 0.0011 0.0010 0.0010 
[23] 0.0010 0.0010 0.0010 0.0009 0.0009 0.0008 0.0008 0.0007 0.0007
\end{CodeOutput}
\end{CodeChunk}
Similarly, 31 standard errors for each simple unbiased coverage estimate defined in (\ref{svar}) are
\begin{CodeChunk}
\begin{CodeInput}
R> pcv$se.coverageS
\end{CodeInput}
\begin{CodeOutput}
 [1] 0.0070 0.0062 0.0063 0.0063 0.0069 0.0072 0.0062 0.0067 0.0062 0.0065 0.0066
[12] 0.0072 0.0066 0.0066 0.0062 0.0058 0.0066 0.0068 0.0062 0.0065 0.0063 0.0066 
[23] 0.0059 0.0062 0.0072 0.0074 0.0071 0.0062 0.0075 0.0072 0.0065
\end{CodeOutput}
\end{CodeChunk}

Taking the first hospital as an example, the variance estimate of RB unbiased coverage estimate is about 19 times smaller than that of simple one. It means that 1,000 RB unbiased coverage estimates are as precise as 17,000 simple unbiased coverage estimates in terms of estimating true coverage probability for the first hospital, $p_{cov, 1}$.

@@

The Education Testing Service (ETS) conducted randomized experiments in eight separate schools (groups) to test whether students (units) SAT scores are effected by coaching. The dataset contains the estimated coaching effects on SAT scores ($y_{j}, j=1, \ldots, 8$) and standard errors ($se_{j}, j=1, \ldots, 8$) of the eight schools \citep{1981}.
\begin{CodeChunk}
\begin{CodeInput}
R> y  <- c(12, -3, 28,  7,  1,  8, 18, -1)
R> se <- c(18, 16, 15, 11, 11, 10, 10,  9)
\end{CodeInput}
\end{CodeChunk}
or
\begin{CodeChunk}
\begin{CodeInput}
R> data(schools)
R> y  <- schools$y
R> se <- schools$se
\end{CodeInput}
\end{CodeChunk}

@@

\begin{figure}[h] 
\begin{center}
\includegraphics[width = 3.2in]{school1.png}
\caption{Shrinkage plot and 95\% interval plot for 8 schools}
\label{fig:8schoolsplot}
\end{center}
\end{figure}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> z <- c(18, 17, 16, 15, 14, 14, 13, 12, 11, 11, 10, 10, 10, 10, 10,  9,  8,  7)
R> n <- c(45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45)
R> x <- c( 1,  1,  1,  1,  1,  0,  0,  0,  0,  1,  0,  0,  0,  1,  1,  0,  0,  0) 
\end{CodeInput}
\end{CodeChunk}
or
\begin{CodeChunk}
\begin{CodeInput}
R> data(baseball)
R> z <- baseball$Hits
R> n <- baseball$At.Bats
R> x <- ifelse(baseball$Position == "fielder", 1, 0)
\end{CodeInput}
\end{CodeChunk}
%$
Conditioning on the true batting average for each player we assume that the at-bats are independent and therefore, $z_{j}\vert p_{j}\stackrel{ind}{\sim} \textrm{Binomial}(45, p_{j}), ~j=1, \ldots, 18$. Our goal is to obtain point and interval estimates of the true batting average, $p_{j}$, for each player, whilst considering the additional information on whether the player is an outfielder or not. \code{gbp} provides a way to incorporate such covariate information seamlessly into the second-level hierarchy such that information is shared and regression towards the mean (RTTM) occurs within outfielders and non-outfielders separately. %For reference, we can also assume the Normal distribution using $y_{j}=z_{j} / n_{j}$ and $V_{j}=\bar{y}(1-\bar{y})/n$, where $\bar{y}=\sum_{j} z_{j} / \sum_{j} n_{j}$, $j=1,\ldots, 18$.

@@

\begin{CodeChunk}
\begin{CodeInput}
R> plot(b)
\end{CodeInput}
\end{CodeChunk}
\begin{figure}[h]
\begin{center}
\includegraphics[width = 3.2in]{baseball1.png}
\caption{Shrinkage plot and 95\% interval plot for 18 baseball players}
\label{fig:baseball}
\end{center}
\end{figure}

@@

The Rao-Blackwellized unbiased coverage estimates and their standard errors for each player follow.
\begin{CodeChunk}
\begin{CodeInput}
R> bcv$coverageRB
\end{CodeInput}
%$
\begin{CodeOutput}
 [1] 0.971 0.973 0.972 0.972 0.970 0.973 0.973 0.974 0.973 0.973 0.971 0.973 
[13] 0.973 0.972 0.972 0.971 0.973 0.971
\end{CodeOutput}
\end{CodeChunk}
\begin{CodeChunk}
\begin{CodeInput}
R> bcv$se.coverageRB
\end{CodeInput}
\begin{CodeOutput}
 [1] 0.0015 0.0012 0.0013 0.0014 0.0016 0.0010 0.0012 0.0010 0.0010 0.0013 
[11] 0.0015 0.0013 0.0019 0.0013 0.0014 0.0015 0.0011 0.0014
\end{CodeOutput}
\end{CodeChunk}
